GRM_fourway

### 2020-04-04 R/GRM_kinship_for_fourway.Rmd
Find GRM (kinship) for fourway cross individuals for Li

---
title: "Make Kinship Van Raden"
author: "Alice MacQueen"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
```

@@

In the GAPIT numerical object, columns are SNPs and rows are individuals. It's the same for bigsnpr objects, luckily.

# Making a kinship matrix using the van Raden method.

If you don't need to subset by individuals or SNPs, you can use this function as-is. Otherwise see the code in the next section for options for subsetting.

```{r}
snp <- snp_attach("Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds")

pvdiv_kinship_van_raden <- function(snp){
  G <- snp$genotypes
  nInd <- snp$genotypes$nrow
  # Centered (mean of rows subtracted) transposed crossproduct of snp file.
  K <- big_tcrossprodSelf(G, fun.scaling = big_scale(center = TRUE, 
                                                                 scale = FALSE))

  #Extract diagonals
  i = 1:nInd
  j=(i-1)*nInd
  index=i+j
  d=K[index]
  DL=min(d)
  DU=max(d)
  floor=min(K[i,i])

  K=(K[i,i]-floor)/(DL-floor)
  MD=(DU-floor)/(DL-floor)

  rownames(K) <- snp$fam$sample.ID
  colnames(K) <- snp$fam$sample.ID

  if(MD>2){
    K[index]=K[index]/(MD-1)+1
  }
  return(K)
}
```

# Subset to just single ecotypes, and to just single genetic subpopulations.

This method could be used to remove any individuals that don't meet specific criteria.

```{r}
metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")

upeco <- metadata[which(metadata$ECOTYPE_NNET == "Upland"),]$PLANT_ID
loeco <- metadata[which(metadata$ECOTYPE_NNET == "Lowland"),]$PLANT_ID
coeco <- metadata[which(metadata$ECOTYPE_NNET == "Coastal"),]$PLANT_ID

mgeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Midwest"),]$PLANT_ID
ageno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Atlantic"),]$PLANT_ID
ggeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Gulf"),]$PLANT_ID

upland_eco <- which(snp$fam$sample.ID %in% upeco)
lowland_eco <- which(snp$fam$sample.ID %in% loeco)
coastal_eco <- which(snp$fam$sample.ID %in% coeco)
midwest_geno <- which(snp$fam$sample.ID %in% mgeno)
atlantic_geno <- which(snp$fam$sample.ID %in% ageno)
gulf_geno <- which(snp$fam$sample.ID %in% ggeno)

subsets <- list(upland_eco = upland_eco, lowland_eco = lowland_eco, coastal_eco = coastal_eco, midwest_geno = midwest_geno, atlantic_geno = atlantic_geno, gulf_geno = gulf_geno)

for(q in seq_along(subsets)){
    ind.row <- subsets[[q]]
    G <- snp$genotypes
    nInd <- length(ind.row)
    # Centered (mean of rows subtracted) transposed crossproduct of snp file.
    K <- big_tcrossprodSelf(G, ind.row = ind.row, 
                            fun.scaling = big_scale(center = TRUE, 
                                                    scale = FALSE))

    #Extract diagonals
    i = 1:nInd
    j=(i-1)*nInd
    index=i+j
    d=K[index]
    DL=min(d)
    DU=max(d)
    floor=min(K[i,i])

    K2=(K[i,i]-floor)/(DL-floor)
    MD=(DU-floor)/(DL-floor)
    # Name rows and columns
    rownames(K2) <- snp$fam$sample.ID[ind.row]
    colnames(K2) <- snp$fam$sample.ID[ind.row]

    if(MD>2){
      K2[index]=K2[index]/(MD-1)+1
    }

  write_rds(K2, paste0("Kinship_van_Raden_method_", nInd, "_individuals_", names(subsets)[q], "type.rds")) 
}
```

Other bigstatsr commands that I didn't end up using.
`snpSum <- big_colstats(G)# GAPIT: apply(snps,2,mean) Take the mean over columns.`
`snpMean <- snpSum$sum / nInd # bigsnpr has a function that can center like this.`

@@

>>>>

---
title: "GRM_fourway"
author: "Alice MacQueen"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)
```

@@

```{r}
# paste0(rep(c(1:9),each = 2), rep(c("K", "N"), 9))

# str(cross$geno$`2N`$data)

geno_df <- cross$geno[[1]]$data

for(i in 2:18){
geno_df <- geno_df %>%
  cbind(cross$geno[[i]]$data)
}

# geno_df[1:20,1000:1005]
# length(which(!is.na(geno_df)))
# 50666/6170584 # less than 1% missing data, but it majorly fucks up the kinship matrix.

geno_df2 <- geno_df %>%
  as_tibble() %>%
  mutate_if(is.numeric, list(~ifelse(is.na(.), sample(1:4, 1), .))) # replace NA's with a random value from 1 to 4.

snps <- as.matrix(geno_df2)
# snps[1:20,1000:1005]
```

@@

```{r}
K[1:10,1:5]
library(switchgrassGWAS)

mash_plot_pairwise_sharing(corrmatrix = K[1:30,1:30])

write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_5250g_fourway_cross.rds")
```
